### 2015-01-04 mcmc.tex
methods

\paragraph{Tempering:}
... HOGG

\paragraph{Nested:}
Similar to tempering are a class of methods called ``nested''
samplings, which also smoothly change the target distribution.
In the nested case, instead of increasing the temperature, the
samplings are of a censored version of the prior, censored by the
value of the likelihoood function.

\paragraph{Reversible-jump:}
Finally, there are some problems that \emph{cannot be done} without
additional proposal technology...

>>>>

\paragraph{Tempering:}
A huge problem for MCMC samplers---which make local moves based on
local positioning in the parameter space---is that there can be
multiple modes in the function $f(\pars)$, separated by valleys of
zero or near-zero density.
A vanilla MCMC method cannot easily cross such valleys with local
moves.
The idea behind simulated tempering and related methods is to ``smooth
out'' the function $f(\pars)$ or reduce its dynamic range, to make the
peaks less tall and the valleys less deep, to permit the MCMC random
walk to cross the valleys.
The standard method is to introduce a ``temperature'' variable and,
while sampling, take the likelihood function to the power of the
inverse temperature (or, if working in the logarithm, multiply the log
likelihood by the inverse temperature).
When the temperature is high, this reduces the influence of the
likelihood function relative to the prior pdf, and (assuming that the
prior pdf is easy to sample) makes movement in the parameter space
easier.
At each step, when the temperature has moved away from unity, the
distribution being sampled is not exactly $f(\pars)$, but the results
of the sampling are only kept for those samples taken at unit
temperature; these in the end make a fair sampling.

\paragraph{Nested:}
Similar to tempering-like methods are a class of methods called ``nested''
samplers, which also smoothly change the target distribution away from $f(\pars)$.
In the nested case, instead of increasing the temperature, the
samplings are of a censored version of the prior, censored by the
value of the likelihoood function.
When the likelihood censoring is strong, only the most high-likelihood
parts of the prior get sampled; when the likelihood censoring is weak,
almost all of the prior gets sampled.
The idea behind nested sampling---like tempering---is that it is
designed to be a good method for exploring the full parameter space or
searching for all of the modes of the posterior pdf.
It is also designed such that it produces not just a sampling but also
an estimate of the integral $Z$ of the likelihood times the prior
(the Bayes factor or fully marginalized likelihood).

\paragraph{Reversible-jump:}
Finally, there are some extreme problems in which it is not just hard
to sample in the parameter space, but the parameter space itself is
not of fixed dimension.
For example, if you are modeling an astronomical image with a
collection of stars, and you don't know how many stars to use; in this
case the number of stars itself is a parameter, so the number of
parameters is itself a function of the parameters.\note{HOGG Cite
  Brewer et al}
In these cases you can only use certain kinds of samplers (though M--H
MCMC is one you can choose), and, in addition, you need to make
special kinds of proposals that permit the system to jump from one
parameter space to another that has different dimensionality.
In these cases, the concept of detailed balance becomes non-trivial,
but there are criteria for creating reversible-jump proposals between
the parameter spaces.
These methods are best built and operated under the supervision of a
trained professional.
